# Data {#ch:Data}

Each of the six federal elections are modelled with a cross-sectional data set containing information about each electorate in that election. This consists of the two-party preferred vote, which is the response variable, and socio-demographic variables which are explanatory variables that characterise each electorate. 

In this section, we specify the sources for the two-party preferred vote and socio-demographic variables, detail the preliminary preparations of the raw data, discuss why imputation is required to obtain the desired cross-sections for each election, and illustrate our chosen imputation algorithm. Additionally, we explain why standardizing the explanatory variables is important in order to compare models across elections.

## Preliminary

The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections. The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives. Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate. The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered. This is recorded as a percentage preference in favour of the Liberal party.

A map of electoral boundaries in each election is also provided by the AEC in geographical information system (GIS) format.

For each election the two-party preferred vote is obtained for all 150 electorates, with the exception of 2001. The electoral two-party preferred vote is not published for the 2001 election. Instead, we calculate it from the two-candidate preferred vote for the 141 electorates that had Labor and Liberal as the two most popular candidates, leaving nine electorates without a two-party preferred vote recorded. These 9 electorates are omitted from the subsequent analysis.

Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census, and publishes aggregated information (including GIS maps of the aggregated regions). We choose the aggregate level of Commonwealth Electoral Divisions, which is the ABS approximation of electorates at the time of the Census. From this aggregate information, 63 socio-demographic variables are computed for each of the Commonwealth Electoral Divisions. A full description of these variables can be found in Appendix \@ref(a:describeSD). From here on, Commonwealth Electoral Divisions will be referred to as *Census divisions*, so that they are not confused with electorates.

At this stage, we have obtained the electoral two-party preferred votes for the six elections, and socio-demographic variables for Census divisions in the four Censuses. In order to model each election, the socio-demographic variables need to be linked with the electorates in each election. Elections and Censuses are not contemporaneous, nor are the boundaries of electorates guaranteed to align with a set of Census divisions - as the AEC redistributes electoral boundaries. Therefore, a method of imputation is required to obtain the desired cross-sections for each election.

The next steps are to determine how the Census division and election boundary information differ in both temporal and spatial dimensions, and to devise an imputation procedure.

## Understanding the difference

First, consider the temporal difference between election data (two-party preferred vote) and Census data. Six elections and four Censuses were conducted from 2001 to 2016, but only in 2001 and 2016 did both events occur (see Figure \@ref(fig:timeline)). In order to match socio-demographic information to the other four elections, we need to interpolate between Censuses in some way. 

```{r timeline, fig.cap="Timeline of Australian elections and Censuses", out.width = "90%", fig.align="center"}
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(election,census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "census", year + 0.1, year))


ggplot(timeline, aes(x=year_vis, y=measure)) + geom_point(aes(color=as.factor(event),shape=as.factor(event)),size = 4) + labs(color = "Event", shape = "Event") + geom_hline(yintercept=0, size =1, color = 'grey', show.legend = FALSE) + geom_segment(x=2001,xend=2001, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2004,xend=2004, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2007,xend=2007, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2010,xend=2010, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2013,xend=2013, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2016,xend=2016, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2001.1,xend=2001.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2006.1,xend=2006.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2011.1,xend=2011.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2016.1,xend=2016.1, y=1.4, yend=0, color = '#F8766D') + labs(x="Year") + theme(text=element_text(size=8), axis.text.x=element_text(size=6),axis.text.y=element_blank(), axis.title.y=element_blank(), plot.background=element_blank(), axis.ticks.y = element_blank(), panel.background = element_blank()) + scale_x_discrete(limits = c(2001:2016)) + coord_fixed(ratio = 2)
```

Recall that the Census is aggregated to Census divisions, and the two-party preferred vote is aggregated to the electorate level. Electorate boundaries are redrawn regularly by the AEC, and so Census divisions match the electorate boundaries at the time of the Census. Therefore, when a Census is conducted in an election year, such as in 2001 and 2016, the electoral boundaries match the Census divisions. However, for elections in between (2004, 2007, 2010 and 2013), there is no match between Census divisions and electoral boundaries. Therefore, to impute socio-demographic variables for each electorate in these elections, the spatial differences in Census and election data must also be taken into account.

## Spatio-temporal imputation

Now that the spatial and temporal differences between election and Census data have been explained, we outline the chosen method of spatio-temporal imputation. This procedure estimates electoral socio-demographic variables for an election.

The essence of the algorithm is to estimate the socio-demographic composition of electorates had their boundary been in place at the time of the two nearest Censuses, one before and one after, and then to linearly interpolate between these time points. For example, the 2013 electoral boundaries are estimated at the time of the 2011 and 2016 Censuses. This algorithm predominantly uses tools from the `sp`, `rgeos` and `dplyr` packages, and is formulated from the piecewise approximation method of areal imputation in @Goodchild1993. 

We now detail the steps of the algorithm, followed by an example.

**Algorithm: Spatial imputation**

What is needed: 

- GIS maps of electoral boundaries and Census division boundaries for the nearest Census before and after the election. 
- Socio-demographic variables for the Census divisions in these two Censuses.

**1**. Select the Census after the election. 

**2**. Select an electorate.

**3**. Determine which Census divisions intersect with the given electorate. (This is done by applying `gIntersects` between the GIS polygon representing the electorate and each of the Census division polygons.)

**4**. For each intersecting Census division, compute the area of overlap with the electorate and determine the proportion of the Census division that is consumed by this overlap. (`gIntersection` determines the area of intersection between two polygons, and `gArea` computes the area of a given polygon.)

Let $\alpha_{e,c} \in [0,1]$ be the proportion of Census division $c$ that is contained in the overlapping region with electorate $e$.

**5**. Estimate the population in each overlapping area by assuming populations of Census divisions are uniformly distributed.

Let $\theta_{c}$ be the total population in Census division $c$, and $p_{e,c}$ denote the estimated overlap population size between electorate $e$ and Census division $c$. Then

$$\hat{p}_{e,c} = \alpha_{e,c} \cdot \theta_{c}.$$

**6**. For each socio-demographic variable, compute a weighted average of the variable's value from each of the intersecting divisions, using the estimated populations as proportionate weights. The result is the estimate of the socio-demographic variable for the electorate at the time of the Census. Repeat this process for all socio-demographic variables.

Let $\lambda_{e,t}$ be the value of a given socio-demographic variable for electorate $e$ in year $t$, and $\delta_{c,t}$ be the value of the same variable for Census division $c$. Now add a subscript $t$ (to denote the relevant year) to the estimated overlap population $\hat{p}_{e,c,t}$ between electorate $e$ and Census division $c$. 

$$\hat{\lambda}_{e,t} = \frac{\sum_{c=1}^{150}  \hat{p}_{e,c,t}*\delta_{c,t}}{\sum_{c=1}^{150}  \hat{p}_{e,c,t}}$$
**7**. Repeat steps 3-6 for all of the 150 electorates, using the same Census year.

**8**. Repeat steps 3-7 using the other Census year relevant to the election. After this step, all socio-demographic variables for the electoral boundaries will have been estimated for both Censuses.

**9**. Linearly interpolate between consecutive Census years for each electorate to impute the socio-demographic profile for the election year.

Let $t_1$ be the year of the Census prior to the election, $s$ be the year of the election, and $t_2$ be the year of the Census after the election.

$$\hat{\lambda}_{e,s} = \frac{s-t_1}{t_2-t_1} \cdot \hat{\lambda}_{e,t_1} + \frac{t_2-s}{t_2-t_1} \cdot \hat{\lambda}_{e,t_2}$$

**Example: electorate of Hume in the 2013 election.**

To illustrate the spacio-temporal algorithm detailed above, consider the imputation of a socio-demographic variable for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election. Figure \@ref(fig:hume13) shows this region amongst other NSW electorates.

```{r hume13, fig.cap = "Electoral boundaries for 2013, with the electoral boundary for Hume, shown in purple."}
# 2013
hume_only <- map13 %>% filter(Election_Division == "HUME")
min_lat = min(hume_only$lat - 0.5)
max_lat = max(hume_only$lat + 0.5)
min_long = min(hume_only$long - 0.5)
max_long = max(hume_only$long + 0.5)

map_hume_surrounds_13 <- map13 %>% 
  group_by(Election_Division) %>% 
  mutate(lower_lat = min(lat),
         upper_lat = max(lat),
         lower_long = min(long),
         upper_long = max(long)) %>% 
  ungroup() %>% 
  filter(lower_lat < max_lat, upper_lat > min_lat, lower_long < max_long, upper_long > min_long,
         !Election_Division %in% c("PARKES", "FARRER"))

map_hume_surrounds_13  %>% 
ggplot() +
  geom_polygon(aes(x=long, y=lat, group=group,
               fill=factor(Election_Division == "HUME")), colour="black", size = 0.2, alpha = 0.4) +
  scale_fill_manual(values = c("white", "purple"), name = "", labels = c("Other", "Hume")) +
  lims(x = c(144,152), y=c(-38,-32)) + 
  coord_equal() + 
  theme_bw() + 
  theme(legend.position="bottom", plot.title = element_text(face = "bold", size = 12, hjust = 0.5), 
        axis.title = element_text(size = 8), axis.text = element_text(size = 8), legend.text = element_text(size = 8),
        legend.key.size = unit(0.5,"cm")) + 
  labs(x="Longitude", y="Latitude", title = "2013")
```

The Censuses neighbouring the 2013 election are those in 2011 and 2016. Let's start with drawing the boundary for Hume from the 2013 election over the Census divisions from the 2016 Census. Figure \@ref(fig:hume16) shows the same region of NSW as in Figure \@ref(fig:hume13), but now with the with the 2013 electorate of Hume shaded in purple. This purple region is the area whose socio-demographic variables are to be imputed.

There are many Census divisions that intersect with the Hume (purple) region, these include the Census divisions of Riverina, Eden-Monaro and Hume which we refer to as CD1, CD2 and CD3. There are also tiny intersecting areas with Fenner, Calare, Gilmore and Whitlam, but we will ignore these intersections for the purpose of this illustration. Figure \@ref(fig:hume16int) shows these three major intersecting Census divisions in colour.

```{r hume16, fig.cap = "Census division boundaries for 2016, with the 2013 electoral boundary for Hume, shown in purple."}
# Surrounding electorates 2011 and 2016
hume_elec <- map13 %>% filter(Election_Division == "HUME")
min_lat = min(hume_elec$lat - 0.5)
max_lat = max(hume_elec$lat + 0.5)
min_long = min(hume_elec$long - 0.5)
max_long = max(hume_elec$long + 0.5)

map_hume_impute <- map11 %>% mutate(part = "2011") %>% 
  bind_rows(hume_elec %>% mutate(part = "2011"),
            hume_elec %>% mutate(part = "2016"),
            map16 %>% mutate(part = "2016")) %>%
  group_by(Election_Division, year) %>% 
  mutate(lower_lat = min(lat),
         upper_lat = max(lat),
         lower_long = min(long),
         upper_long = max(long)) %>% 
  ungroup() %>% 
  filter(lower_lat < max_lat, upper_lat > min_lat, lower_long < max_long, upper_long > min_long,
         !Election_Division %in% c("PARKES", "FARRER"))
 
## Faceted - 2016 only
ggplot(data=map_hume_impute %>% 
         filter(year != "2011") %>% 
         filter(year != "2013")) +
  geom_polygon(aes(x=long, y=lat, group=group), colour="black", fill = "white", size = 0.2) +
  geom_polygon(aes(x=long, y=lat, group=group,
               fill=factor(Election_Division == "HUME" & year == "2013")), colour=NA, fill = "purple", size = 0.2, alpha = 0.4,
               data = map_hume_impute %>% filter(year == "2013")) +
  scale_fill_manual(values = c("white", "purple")) +
  lims(x = c(146,152), y=c(-38,-32)) + 
  coord_equal() + 
  theme_bw() + 
  theme(legend.position="bottom", plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title = element_text(size = 8), axis.text = element_text(size = 8)) + 
  #facet_wrap(~part) + 
  labs(x="Longitude", y="Latitude", title = "2016") 
```

```{r hume16int, fig.cap = "Census division boundaries for 2016, with the 2013 electorate boundary for Hume, shown in purple, and with intersecting Census divisions highlighted."}
## Faceted - 2016 only
## Includes shaded regions for intersecting divisions
plot_map_hume_impute <- map_hume_impute %>%
         filter(year != "2011") %>% 
         filter(year != "2013") %>% 
         mutate(intersecting = ifelse(Election_Division == "RIVERINA", 2,
                                      ifelse(Election_Division == "EDEN-MONARO", 3,
                                             ifelse(Election_Division == "HUME" & year == "2016", 1, 0))))

ggplot(data=plot_map_hume_impute) +
  geom_polygon(aes(x=long, y=lat, group=group, fill = factor(intersecting)), colour="black", size = 0.2, alpha = 0.2) +
  geom_polygon(aes(x=long, y=lat, group=group), colour=NA, fill = "purple", size = 0.2, alpha = 0.4,
               data = map_hume_impute %>% filter(year == "2013")) +
  geom_polygon(aes(x=long, y=lat, group=group), colour="black", fill = "white", size = 0.2, alpha = 1,
               data = plot_map_hume_impute %>% filter(Election_Division %in% c("CANBERRA", "FENNER"))) +
  scale_fill_manual(values = c("white", "red", "yellow", "green", "purple"), labels = c("Other", "CD1", "CD2", "CD3", "Area to be estimated")) +
  lims(x = c(146,152), y=c(-38,-32)) + 
  coord_equal() + 
  theme_bw() + 
  theme(legend.position="right", plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title = element_text(size = 8), axis.text = element_text(size = 8)) + 
  #facet_wrap(~part) + 
  labs(x="Longitude", y="Latitude", title = "2016", fill = "Census divisions")
```

The 2013 electoral boundary for Hume (purple) covers  $90.58 \%$ of CD1, $24.89 \%$ of CD2 and $9.96 \%$ of CD3. Having obtained these percentage, we now estimate the population in these intersections via the assumption of uniformly distributed populations within each Census division. For example, the population of CD2 is $155,793$, therefore $38,780$ people are estimated to be in the region of CD1 that intersects with Hume. 

```{r, results = 'asis'}
## Table
Mapping_aec13 %>% 
  filter(Election_Division == "Hume", Year == "2016", Percent_Elec_Composition > 0.05) %>% 
  select(Census_Division, Year, Percent_Cens_Composition) %>% 
  mutate(Percentage = paste0(round(Percent_Cens_Composition*100,2), "%")) %>% 
  rename(Division = Census_Division) %>% 
  mutate(Division = toupper(Division)) %>% 
  left_join(abs_all %>% 
              filter(Electorate %in% c("HUME", "RIVERINA", "EDEN-MONARO"), year == 2016) %>% 
              select(Population, Electorate) %>% 
              rename(Division = Electorate), by = "Division") %>% 
  mutate(`Estimated Population Allocated to Hume` = (Population*Percent_Cens_Composition) %>% round(0) %>% as.character(),
         Population = Population %>% as.character()) %>% 
  select(-Percent_Cens_Composition) %>% 
  rename(`Population in Division` = Population) %>%
  mutate(intersecting = ifelse(Division == "RIVERINA", "CD2",
                                      ifelse(Division == "EDEN-MONARO", "CD3",
                                             ifelse(Division == "HUME", "CD1", 0)))) %>% 
  select(intersecting, Year, Percentage, `Population in Division`, `Estimated Population Allocated to Hume`) %>% 
  rename(`Census Division` = intersecting) %>% 
  xtable() %>% 
  print(comment=FALSE, digits=NULL, size = "\\small")
```

Now we compute a weighted average of each socio-demographic variable from the intersecting divisions, using the estimated populations as weights. Recall that $\delta_{c,t}$ represents a given socio-demographic variable for Census division $c$ at time $t$, and that $\hat{\lambda}_{e,t}$ is our estimate of the variable for electorate $e$ at time $t$.

$$\hat{\lambda}_{Hume, 2016} = \frac{136,458 \cdot \delta_{CD1, 2016} +  38,780 \cdot \delta_{CD2, 2016} +  14,691 \cdot \delta_{CD3, 2016}}{136,458 + 38,780 + 14,691}.$$
So here $\hat{\lambda}_{Hume, 2016}$ corresponds to the value of the variable estimated for the 2013 Hume electorate at the time of the 2016 Census. 

For example, consider the percentage of the population who are Australian citizens. In 2016, CD1 had 90.0%, CD2 had 89.1% and CD3 had 88.0% Australian citizen. Then the estimated percentage of Australian citizens in the 2013 electorate of Hume, at the time of the 2016 Census is

$$\hat{\lambda}_{Hume, 2016} = \frac{136,458 \cdot 90.0 \% +  38,780 \cdot 89.1 \% +  14,691 \cdot 88.0 \%}{136,458 + 38,780 + 14,691} = 89.66 \% .$$

This weighted average calculation is repeated for all 63 socio-demographic variables. Then, the entire process is repeated using the 2011 Census, where we find $\hat{\lambda}_{Hume, 2011} = 89.58 \%$ Australian citizens.

Moving now to temporal interpolation, we note that the 2013 election sits two years after the 2011 Census, and three years before the 2016 Census. Therefore

\begin{eqnarray*}
\hat{\lambda}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{\lambda}_{Hume,2011} + \frac{2}{5} \cdot \hat{\lambda}_{Hume,2016} \\ 
& = & \frac{3}{5} \cdot 89.58 + \frac{2}{5} \cdot 89.66 \% \\ 
& = & 89.61
\end{eqnarray*}


Therefore, the estimated percentage of Australian citizens in Hume at the time of the 2013 election is 89.61%. This same approach is applied for each of the 63 socio-demographic variables.

Now that electoral socio-demographics are imputed for each election, we consider why standardizing these explanatory variables is an important transformation before modelling each election

## Standardizing the explanatory variables

For many of the socio-demographic variables, both the sample mean and variance change over the years. Consider median rental price in each electorate, for example. In 2001, the median rental price amongst electorates had a mean of \$161.83 and a standard deviation of \$50.89, whereas in 2016 the mean was \$231.49 and the corresponding standard deviation was \$61.88. These figures have been adjusted for inflation using a 2001 index. This means that a \$200 dollar rental price in 2001 was expensive, whereas it was cheap in 2016. Therefore, to make comparisons regarding the effect of rental prices on the two-party preferred vote across 2001 and 2016 elections, rental prices are centered to have mean 0, and standardized to have variance 1. This is done for all socio-demographic variables within each election year, so that data scales are consistent across years.

The values for an electorate can therefore be interpreted as a score relative to the other electorates in that election. For example, a value of $\text{AusCitizen} = 1$ corresponds with an electorate having a proportion of Australian citizens that is one standard deviation above the average in that election.

## Changing socio-demographic of Australia

Once a complete collection of socio-demographic variables is available for each electorate in all six elections, violin plots can be used to analyse how electoral socio-demographic characteristics have changed over time. Violin plots depict the distribution as a density plot reflected in the vertical axis. Trends can be identified over this 17 year period by placing plots from censecutive elections together in the same frame. Changes in age, religion and education are explored in the sections below. 

#### The population is ageing

It is well documented that Australia's population is ageing [@Rich08], and this is reflected in the plots of median age (Figure \@ref(fig:vis-age)). The maximum median age amongst electorates increases across elections, from 41 years in 2001 to 50 years old in 2016. Interestingly, the electorate of Lyne (coastal NSW) has the highest median age in every year. Conversely, the minimum median age has barely changed, with electorates like Melbourne (VIC) having a median age hovering around 31 years. Some electorates have moved in the rank of median age over the years, with electorates like Mayo (SA) moving from having an typical age profile (36 years in 2001), to being one of the older areas (46 years in 2016).

```{r vis-age, out.height='25%', fig.cap="Distribution of median age across Australian electorates, by election year."}
profiles %>% 
  ggplot(aes(x=year, y=MedianAge, fill=year)) + 
  geom_violin() + 
  labs(x = "Election year", y = "Median age of electorate (years)") +
  guides(fill = F) +
  theme(axis.text.x = element_text(size = 12, angle = 60, vjust = 0.5), axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12), aspect.ratio = 0.5)
```


```{r, include = FALSE}
#profiles %>% filter(year %in% c("2001","2016")) %>% select(MedianAge, Election_Division, year) %>% spread(year, MedianAge) %>% mutate(change = `2016` - `2001`)
```

#### Religion: a thing of the past?

Socially "progressive" movements continue to gather momentum all over the world, and Australia is no different, with many people dissociating from traditional religious beliefs and values. The frequency of individuals identifying themselves with a religion has declined over the years, as shown in Figure \@ref(fig:vis-relig). This trend is observed across all electorates, as the minimum value has increased. In 2016, having any religious affiliation would make you a minority in some electorates! 

```{r vis-relig, out.height='25%', fig.cap="Distribution of no religion across Australian electorates, by election year"}
profiles %>% 
  ggplot(aes(x=year, y=NoReligion, fill=year)) + 
  geom_violin() + 
  labs(x = "Election year", y = "Percentage of electorate not identifying \nwith a religion (%)") +
  guides(fill = F) +
  theme(axis.text.x = element_text(size = 12, angle = 60, vjust = 0.5), axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 12), aspect.ratio = 0.5)
```

#### Investing in education

Improvement in educational outcomes are evident across all of Australia. Completion rates for high school, Bachelor's degrees and postgraduate degrees have increased for all electorates (see Figure \@ref(fig:vis-educ)). In 2001, all three measures appear positively skewed, and over the years this skewness has reduced and the distributions have flattened. This indicates that some areas with lower levels of education have progressed at a faster rate than the rest of the country.

```{r vis-educ, out.width='100%', fig.cap="Distribution of educational qualification across Australian electorates, by election year."}
profiles %>% 
  select(year, Bachelor, HighSchool, Postgraduate) %>% 
  gather("metric", "value", -year) %>% 
  ggplot(aes(x=year, y=value, fill=year)) + 
  geom_violin() + 
  facet_wrap(~metric, scales = "free_y") +
  labs(x = "Election year", y = "Percentage of electorate population \nwith qualification (%)") +
  guides(fill = F) +
  theme(axis.text.x = element_text(size = 8, angle = 60, vjust = 0.5), axis.text.y = element_text(size = 8), axis.title = element_text(size = 8), aspect.ratio = 0.75)
```

Illustrating the changes in these three areas demonstrates that socio-demographic characteristics of electorates have evolved over the years.

In this section we have detailed the steps of data collection, preliminary transformations and imputation that result in a cross-sectional data set of electoral socio-demographics and two-party preferred vote for each of the six elections. Violin plots summarising the full set of socio-demographic variables can be found in Appendix \@ref(a:summary). Socio-demographic (explanatory) variables are standardized so that the data scales are consistent across years, which allows for comparison of effects across elections. The next steps are to determine what econometric model to pursue, and how to deal with the problem of having a small number of observations each election ($n=150$), relative to the number of available predictors ($p=63$). This is explored in Section \@ref(ch:Modelling).

