# Data {#ch:Data}

The Census provides information about characteristics of every household in Australia, and is conducted every five years by the Australian Bureau of Statistics (ABS). There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016. The information contained in these collections are used to build the socio-demographic profiles for each electorate, at election time.

Federal elections typically occur every three years, and those of interest are the 2001, 2004, 2007, 2010, 2013 and 2016 elections. We source the two-party preferred vote from the Australian Electoral Commission (AEC), which is an independent federal agency in charge of organising, conducting and supervising federal elections.

All information used in this study is publicly available.

In order to model an election, we need to construct socio-demographic profiles for each electorate, so that it can be paired with the two-party preferred vote. Since the timing of Censuses and elections do not always align (figure \@ref(fig:timeline)), and electorate boundaries change over time, Census profiles for the 2004, 2007, 2010 and 2013 elections must be imputed accounting for both temporal and spatial elements.

```{r timeline, fig.cap="Timeline of Australian elections and Censuses", out.height="30%", out.width = "90%", fig.align="center"}
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(election,census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "census", year + 0.1, year))


ggplot(timeline, aes(x=year_vis, y=measure)) + geom_point(aes(color=as.factor(event),shape=as.factor(event)),size = 4) + labs(color = "Event", shape = "Event") + geom_hline(yintercept=0, size =1, color = 'grey', show.legend = FALSE) + geom_segment(x=2001,xend=2001, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2004,xend=2004, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2007,xend=2007, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2010,xend=2010, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2013,xend=2013, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2016,xend=2016, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2001.1,xend=2001.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2006.1,xend=2006.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2011.1,xend=2011.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2016.1,xend=2016.1, y=1.4, yend=0, color = '#F8766D') + labs(x="Year") + theme(text=element_text(size=8), axis.text.x=element_text(size=6),axis.text.y=element_blank(), axis.title.y=element_blank(), plot.background=element_blank(), axis.ticks.y = element_blank(), panel.background = element_blank()) + scale_x_discrete(limits = c(2001:2016)) + coord_fixed(ratio = 2)
```

## Electorates and their boundaries

Australian Federal elections are determined by the number of seats won by each party in the House of Representatives. Australia is divided into 150 electorates, and each one corresponds with a single seat. The electorate boundaries are determined by population, and to ensure equal representation, the boundaries have to be redrawn regularly by the AEC. Each redistribution typically affects a handful of electorates, with most remaining the same as previously defined.

Since changes are frequently made to electoral boundaries, when it is time for a Census to be conducted, the ABS constructs an approximation to each current electorate and aggregates data at this level. These are called Commonwealth electoral divisions, but will be referred to as *Census divisions*.

What this means is that the electorate boundaries in the election prior to a Census may not match the Census division boundaries used in the Census, which may not match the boundaries in the following election and so on.

This presents a challenge. How does one construct socio-demographic profiles at election time for each electorate, when we cannot directly match that election with a Census?

## Census data

### Source

Census data is downloaded from the ABS website as a collection of Microsoft Excel spreadsheets, with each spreadsheet corresponding to a particular Census division, or a particular question (depending on the format used in that Census year).

In order to convert the information held in these spreadsheets into a summarised table containing selected socio-demographics for each electorate, a series of `R` scripts and `Rmarkdown` files have been created. The output of each file is an `R` data frame object, which tabulates the selected metrics for that Census year.

There was no way to automate this process, and the formats of each Census collection change slightly each year. As such, it has taken a significant amount of time and effort to extract the information from Excel, and wrangle the data into the desired metrics and format. A snippet of the raw data can be found in the Appendix \@ref(fig:excel-demo).

The resultant data frame for each election contains information for each electorate on:

- State
- Population
- Age
- Education
- Employment
- Religious and cultural identity
- Incomes
- Industry of work
- Rental and loan payments
- Citizenship and birthplace
- Language at home
- Relationship status

All metrics are recorded as percentages, representing the percentage of people in that electorate who satisfy the category in question. For example, $AusCitizen$ is the percentage of people in the electorate who are Australian citizens. The exceptions are population and area of each electorate, which are in units of individuals and square kilometers.

A full description of socio-demographic variables in the electorate profiles can be found in the Appendix \@ref(ch:appendix).

It is important to note here that Census divisions is *not* the only level of statistical aggregation available. Census divisions are chosen because they will match electorate boundaries in 2001 and 2016. 

??In an attempt to improve imputation accuracy, we 
and we also extract the same information in post code boundaries (postal areas). Both Census divisions and postal areas could be used in the imputation algorithm, however we find that Census divisions are sufficient for the purpose of this study. This is elaborated in later parts of this thesis.??

##### A note on Census non-response

Like in any survey, non-response bias is a source of potential problems. ABS statements released with each Census assure a high quality of data collection, for all Censuses.

## Elections

Within each electorate, candidates from various political parties run for election to represent that electorate. Voting is compulsory in Australia, and the winning candidate is determined by preferential counts. This means that each person assigns a numbered preference to each candidate, and the winner is determined by receiving a majority of preference votes.

The focus of this study is to differentiate support between the two major partyies - the Australian Labor Party and the Liberal National coalition. The two-party preferred vote is the distribution of preferences where, by convention, comparisons are made only between Labor and Liberal/National candidates. This will be the response variable of interest for each electorate, in each election. The results for each of the six elections are downloaded directly from the AEC website, and formatted to $R$ data.frames, using the same method as described for the Census data.

## Mapping socio-demographic profiles to election times

When a Census is conducted in an election year, the AEC electorate boundaries match the ABS Census divisions, so the two data sources can be linked directly. This is done by for years 2001 and 2016. If the election does not fall on the year a Census is conducted, then socio-demographic information must be imputed for that election.

Any imputation method must account for the:

- Spatial component: electorate boundaries may not match any of the neighbouring Censuses.

- Temporal component: demographics in an area may have changed since the last Census was conducted.

A spatial imputation algorithm using GIS maps is formulated from the piecewise approximation method of areal imputation by @Chooldchild1993. 

The essence of the algorithm is to estimate what a particular electorate would have been comprised of, had it been in place at the time of the nearest Census both before and after the election. This is done predominantly using tools from the `rgeos` package. 

### Spatial imputation explained

To illustrate this algorithm, we will make use of an example - the electorate of Hume (NSW) in the 2013 election. This area is shown in figure \@ref(fig:hume13), shaded in purple with (some of) the neighbouring electorates in New South Wales (NSW).

First the profile at the time of the 2016 Census is estimated, and then that for the 2011 Census - this captures the spatial component. Linearly interpolation over time then accounts for the temporal component. 

*Although here we are focusing on a single electorate, to obtain all electorate socio-demographic profiles for the 2013 election, one would repeat this same process for each electorate individually.*

1. Select the Censuses that occur before and after election time. 

The Censuses from 2011 and 2016 surround the 2013 election.

2. Pick a particular Census year to start with. 

Begin with estimating the electorate boundary had it been in place at the time of the 2016 Census.

3. Select an electorate.

For this example, we have chosen the electorate of Hume.

4. Determine which Census divisions intersect with the electorate of interest.

Figure \@ref(fig:hume16) shows the 2016 Census division boundaries for the same region of NSW as shown in the 2013 electorate boundaries in figure \@ref(fig:hume13), with the 2013 electorate of Hume (purple) superimposed. The purple region is the area we are estimating.

Determine the Census divisions intersect with the electorate region - the divisions of Eden-Monaro, Riverina and Hume. 

*Note that there are also tiny intersecting areas with Fenner, Calare, Gilmore and Whitlam, but we will ignore them for the purpose of this illustration.*

```{r hume16, fig.cap = "Census division boundaries for 2016, with the 2013 electorate boundary for Hume (purple) shown."}
# Surrounding electorates 2011 and 2016
hume_elec <- map13 %>% filter(Election_Division == "HUME")
min_lat = min(hume_elec$lat - 0.5)
max_lat = max(hume_elec$lat + 0.5)
min_long = min(hume_elec$long - 0.5)
max_long = max(hume_elec$long + 0.5)

map_hume_impute <- map11 %>% mutate(part = "2011") %>% 
  bind_rows(hume_elec %>% mutate(part = "2011"),
            hume_elec %>% mutate(part = "2016"),
            map16 %>% mutate(part = "2016")) %>%
  group_by(Election_Division, year) %>% 
  mutate(lower_lat = min(lat),
         upper_lat = max(lat),
         lower_long = min(long),
         upper_long = max(long)) %>% 
  ungroup() %>% 
  filter(lower_lat < max_lat, upper_lat > min_lat, lower_long < max_long, upper_long > min_long,
         !Election_Division %in% c("PARKES", "FARRER"))
 
## Faceted - 2016 only
ggplot(data=map_hume_impute %>% 
         filter(year != "2011") %>% 
         filter(year != "2013")) +
  geom_polygon(aes(x=long, y=lat, group=group), colour="black", fill = "white", size = 0.2) +
  geom_polygon(aes(x=long, y=lat, group=group,
               fill=factor(Election_Division == "HUME" & year == "2013")), colour=NA, fill = "purple", size = 0.2, alpha = 0.4,
               data = map_hume_impute %>% filter(year == "2013")) +
  scale_fill_manual(values = c("white", "purple")) +
  lims(x = c(146,152), y=c(-38,-32)) + 
  coord_equal() + 
  theme_bw() + 
  theme(legend.position="bottom", plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title = element_text(size = 8), axis.text = element_text(size = 8)) + 
  #facet_wrap(~part) + 
  labs(x="Longitude", y="Latitude", title = "2016") 
```

```{r hume16int, fig.cap = "Census division boundaries for 2016, with the 2013 electorate boundary for Hume (purple) and intersecting 2016 divisions highlighted."}
## Faceted - 2016 only
## Includes shaded regions for intersecting divisions
plot_map_hume_impute <- map_hume_impute %>%
         filter(year != "2011") %>% 
         filter(year != "2013") %>% 
         mutate(intersecting = ifelse(Election_Division == "RIVERINA", 1,
                                      ifelse(Election_Division == "EDEN-MONARO", 2,
                                             ifelse(Election_Division == "HUME" & year == "2016", 3, 0))))

ggplot(data=plot_map_hume_impute) +
  geom_polygon(aes(x=long, y=lat, group=group, fill = factor(intersecting)), colour="black", size = 0.2, alpha = 0.2) +
  geom_polygon(aes(x=long, y=lat, group=group), colour=NA, fill = "purple", size = 0.2, alpha = 0.4,
               data = map_hume_impute %>% filter(year == "2013")) +
  geom_polygon(aes(x=long, y=lat, group=group), colour="black", fill = "white", size = 0.2, alpha = 1,
               data = plot_map_hume_impute %>% filter(Election_Division %in% c("CANBERRA", "FENNER"))) +
  scale_fill_manual(values = c("white", "red", "yellow", "green", "purple"), labels = c("Other", "Riverina", "Eden-Monaro", "Hume", "Area to be estimated")) +
  lims(x = c(146,152), y=c(-38,-32)) + 
  coord_equal() + 
  theme_bw() + 
  theme(legend.position="right", plot.title = element_text(face = "bold", size = 12, hjust = 0.5),
        axis.title = element_text(size = 8), axis.text = element_text(size = 8)) + 
  #facet_wrap(~part) + 
  labs(x="Longitude", y="Latitude", title = "2016", fill = "Census divisions")
```


5. Compute the percentage of each Census division that contains in the intersecting region.

The shaded electorate boundary covers $90.58 \%$ of Hume (2016), $24.89 \%$ of Riverina (2016) and $9.96 \%$ of Eden-Monaro (2016).

6. Assuming uniformly distributed populations through the divisions, estimate the population in each of these intersecting regions.

The population of Riverina (2016) is $155,793$, therefore $38,780$ people are estimated to be in the intersecting region of this division. 

```{r, results = 'asis'}
## Table
Mapping_aec13 %>% 
  filter(Election_Division == "Hume", Year == "2016", Percent_Elec_Composition > 0.05) %>% 
  select(Census_Division, Year, Percent_Cens_Composition) %>% 
  mutate(Percentage = paste0(round(Percent_Cens_Composition*100,2), "%")) %>% 
  rename(Division = Census_Division) %>% 
  mutate(Division = toupper(Division)) %>% 
  left_join(abs_all %>% 
              filter(Electorate %in% c("HUME", "RIVERINA", "EDEN-MONARO"), year == 2016) %>% 
              select(Population, Electorate) %>% 
              rename(Division = Electorate), by = "Division") %>% 
  mutate(`Estimated Population` = (Population*Percent_Cens_Composition) %>% round(0) %>% as.character(),
         Population = Population %>% as.character()) %>% 
  select(-Percent_Cens_Composition) %>% 
  xtable() %>% 
  print(comment=FALSE, digits=NULL)
```

7. Compute a weighted average of socio-demographics from the intersecting divisions, using estimated populations as weights.

Let $\lambda_{i}$ is the value of a given socio-demographic for electorate $i$, and $\delta_{j}$ be the value of the same socio-demographic for Census division $j$. Let $p_{j}$ denote the estimated intersection population with Census division $j$.

$$\hat{\lambda}_{i} = \frac{\sum p_j*\lambda_j}{\sum  p_j}$$
Here we have $i = Hume$ and $j = {\text{Hume, Riverina, Eden-Monaro}}$, so:

$$\hat{\lambda}_{Hume} = \frac{136,458 \cdot \delta_{Hume} +  38,780 \cdot \delta_{Riverina} +  14,691 \cdot \delta_{Eden-Monaro}}{136,458 + 38,780 + 14,691}$$

*At this point, we have an estimated 2016 Census profile for each of the 2013 electorate boundaries*.

8. Repeat steps 3-7 for all of the 150 electorates, using the same Census year.

9. Repeat steps 2-8 using the other Census year.

Having already computed estimated for 2016, we would now estimate the electorate profiles for the 2011 Census. 

10. Linearly interpolate between the Census years to impute the socio-demographic division profile for the given election year.

The 2013 election sits two years after the 2011 Census, and three years before the 2016 Census. Take a weighted average of each demographic with respect to time for each division.

Adding a subscript $k$ to $\lambda_{i}$ to denote year, $\lambda_{i,k}$.

$$\hat{\lambda_{i,2013}} = \frac{3}{5} \cdot \hat{\lambda_{i,2011}} + \frac{2}{5} \cdot \hat{\lambda_{i,2016}}$$

## Standardizing the variables

For many of the socio-demographic variables, both mean and variance change over the years. Take median rental prices for example. In 2001, the mean electorate median rental price was $ \$ 161.83$, and standard deivation $ \$ 50.89$, whereas in 2016 the mean was $ \$231.49$ and the standard deviation was $ \$ 61.88$, after adjusting prices for inflation to a 2001 index. Since we are interested in modelling each election separately, and then comparing models across years, we need to center and standardize each variable (to have mean $0$ and variance $1$), within the election year, so that scales are consistent across years.

As a result, we interpret the values for an electorate as a being relative to the other electorates in that election. For example, $AusCitizen_j = 1$ in 2013, means that electorate $j$ has a proportion of Australian citizens 1 standard deviation above the mean.

## A summary of the data

Now that socio-demographic profiles have been imputed for the 2004, 2007, 2010 and 2013 elections, a complete collection of data has been obtained every electorate in every election. The violin plots in figure \@ref(fig:violinunscaled) summarise the socio-demographics that are included in the final models (see the Modelling section for how these variables were constructed and chosen).

```{r violinunscaled}
# Unscaled violin plots go here
load("../Data/Clean/profiles.rda")
profiles <- profiles %>% 
  rename(AdminJobs = ManagerAdminClericalSales,
         ExtractiveJobs = Extractive,
         BornUK = Born_UK,
         BornAsia = Born_Asia,
         BornSEEurope = Born_SE_Europe,
         BornMidEast = Born_MidEast,
         OneParentHouse = OneParent_House)

profiles %>% 
  select(NoReligion, CurrentlyStudying, MedianAge, OneParentHouse, BornUK, BornSEEurope, ExtractiveJobs, AdminJobs, HighSchool, Bachelor, Postgraduate, MedianRent, MedianLoanPay, Perc_LNP, year) %>% 
  filter(!is.na(Perc_LNP)) %>%
  select(-Perc_LNP) %>% 
  gather(key = "metric", value = "value", -year) %>%
  ggplot(aes(x=year, y=value)) +
  geom_violin(aes(fill = year)) + 
  facet_wrap(~metric, scales = "free_y", ncol = 4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, size = 8)) +
  guides(fill = FALSE)
```

```{r violinscaled}
# Standardized violin plots go here
data_mod %>% 
  select(c(superset$varname %>% unlist), Perc_LNP, year) %>% 
  filter(!is.na(Perc_LNP)) %>% 
  gather(key = "metric", value = "value", -year) %>% 
  ggplot(aes(x=year, y=value)) +
  geom_violin(aes(fill = year)) + 
  facet_wrap(~metric, scales = "free", ncol = 4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, size = 8)) +
  guides(fill = FALSE)
```

Summary plots of the excluded variables can be found in appendix X. [[violins of the scaled/unscaled omitted variables]]

### Insights on the changing Australian demographic

-- Edit writing from here: need to adjust everything below to reference the summary graphs above.

Comparing distributions across elections reveals many insights into how the Australian demographic has changed over the 17 year period. Trends can be identified for the entire country and for the spread amongst electorates.

#### As Australians grow old, some stay young

It is well documented that Australia has an ageing population [@Rich08], and this is reflected in the age profile plots shown in Figure \@ref(fig:vis-age). Each of the selected plots displayed (called 'violin plots') show the shape of the distribution of the identified age category (age 15-19, 25-34, 55-65) or median electorate age, by Census year. These are effectively density plots reflected in the vertical axis.

Figure \@ref(fig:vis-age) shows the distribution of median age of each electorate has spread upwards over time, with some electorates in 2016 reaching a median age of 50 years. At the same time, the minimum median age remains at 30 years.

```{r vis-age, fig.asp = 0.2, fig.cap="Age profiles of Australian electorates, by Census year"}
vis_age <- vis_df %>%
  filter(measure %in% c("Age15_19", "Age25_34", "Age55_64", "MedianAge"))

ggplot(vis_age, aes(x=year, y=value, fill=year)) + geom_violin() + facet_wrap(~measure, scales = "free_y", nrow = 1) + labs(x = "Census year", y = "% of electorate") + theme(text=element_text(size=12), axis.text.x=element_text(size=8), axis.text.y=element_text(size=8)) + guides(fill = FALSE)
```

#### Religion: a thing of the past?

Socially "progressive" movements continue to gather momentum all over the world, and as a result, Australia is moving away from traditional religious beliefs and values. The frequency of individuals identifying themselves with a religion has declined over the years, as shown for various religions in figure \@ref(fig:vis-relig). This effect has stretched across (what appears to be) every electorate, as seen in the fourth panel. Having any religious affiliation would make you a minority in some electorates! At the same time, particular electorates maintain relatively large representation of certain otherwise uncommon religious group, as seen by the thin upper tails of the Islam and Judaism metrics.

```{r vis-relig, fig.asp = 0.2, fig.cap="Religion profiles of Australian electorates, by Census year"}
vis_relig <- vis_df %>%
  filter(measure %in% c("NoReligion", "Christianity", "Islam", "Judaism"))

ggplot(vis_relig, aes(x=year, y=value, fill=year)) + geom_violin() + facet_wrap(~measure, scales = "free_y", nrow = 1) + labs(x = "Census year", y = "% of electorate") + theme(text=element_text(size=8), axis.text.x=element_text(size=6), axis.text.y=element_text(size=8))
```

#### Investing in education

Australia has seen improvements in education outcomes across the board, experiencing continual increases in secondary and tertiary completion rates across the years (Figure \ref(fig:vis-educ)). It is encouraging to see that no electorates appear to be lagging behind, as the minimum values increase each year for all levels of education.

```{r vis-educ, fig.asp = 0.2, fig.cap="Education profiles of Australian electorates, by Census year"}
vis_educ <- vis_df %>%
  filter(measure %in% c("HighSchool", "Bachelor", "Postgraduate", "CurrentlyStudying"))

ggplot(vis_educ, aes(x=year, y=value, fill=year)) + geom_violin() + facet_wrap(~measure, scales = "free_y", nrow = 1) + labs(x = "Census year", y = "% of electorate") + theme(text=element_text(size=8),axis.text.x=element_text(size=6),axis.text.y=element_text(size=8))
```

